Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-02-03 Manuscript/manuscript.Rmd
Results review

We evaluated and compared multiple random forest classification algorithms built on domain data extracted from Hi-C datasets for 2 cell lines C = {GM12878, K562}, at 4 different resolutions R = {10 kb, 25 kb, 50 kb, 100 kb}. For each classifier, we compared 3 predictor types P = {OC, OP, Distance} and 5 class balancing techniques B = {None, RUS, ROS, SMOTE, ROSE}. A system produced from the ensemble framework was defined as: E = {c, r, p, b}, where $c \in C, \quad R \in R, \quad p \in P,$ and $b \in$ B. Letting |X| denote the cardinality of each set, we evaluated $|C| \times |R| \times |P| \times |B|$ different ensemble systems, totaling X (=$x \times 4 \times 3 \times 5$) separate models. For each model, we aimed to assess both performance and identify the genomic annotations that were most influential in correctly labeling the genomic bins that overlapped with flanked TAD boundaries. For simplicity, we present the most significant results, with the complete results available as Supplementary Tables and Figures.

We first examined the class imbalance of the different data resolutions (Supplementary Table ). While the total number of genomic bins decreased as resolution increased, the imbalance ratios remained relatively stable. We found that there was an average imbalance ratio of 0.154 (SD=0.022) and 0.137 (SD=0.0181) for cell lines GM12878 and K562 respectively. The low imbalance ratios indicate a wide disparity between majority and minority classes, and therefore suitable resampling techniques are needed.

>>>>

We evaluated and compared multiple random forest classification algorithms built on domain data extracted from Hi-C datasets for 2 cell lines C = {GM12878, K562}, at 4 different resolutions R = {10 kb, 25 kb, 50 kb, 100 kb}. For each classifier, we compared 3 predictor types P = {OC, OP, Distance} and 5 class balancing techniques B = {None, RUS, ROS, SMOTE, ROSE}. A system produced from the ensemble framework was defined as: E = {c, r, p, b}, where $c \in C, \quad R \in R, \quad p \in P,$ and $b \in$ B. Letting |X| denote the cardinality of each set, we evaluated $|C| \times |R| \times |P| \times |B|$ different ensemble systems, totaling X (=$x \times 4 \times 3 \times 5$) separate models. For each model, we aimed to assess both performance and identify the genomic annotations that were most important in correctly labeling the genomic bins that overlapped with flanked TAD boundaries.

We first examined the class imbalance of the different data resolutions (Supplementary Table ). [??? What table?] While the total number of genomic bins decreased as resolution increased, the imbalance ratios remained relatively stable. We found that there was an average imbalance ratio of 0.154 (SD=0.022) and 0.137 (SD=0.0181) for cell lines GM12878 and K562 respectively. [??? At which resolution?] The low imbalance ratios indicate a wide disparity between majority and minority classes, and therefore suitable resampling techniques are needed.
